Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-09 summary.tex
added convective overstability discussion

\subsection{Caveats and outlooks} 
We highlight below some extensions to the present  
linear models in future works:

\emph{Improved thermal evolution.} We assumed non-adiabatic effects
can be represented by a thermal relaxation timescale $t_c$. For 
application to protoplanetary disks we connected $t_c$ to the disk
properties, but one can improve the realism further by generalizing the
opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
\kappa_d(\rho, T)$ as described in \cite{bell94}. %two temperature
% structure  
However, going beyond the framework of thermal
relaxation requires a proper treatment for radiative diffusion, which
involves the Lapacian operator and additional boundary conditions
must be supplemented. We will consider these improvements in a
follow-up study.

\emph{Additional physics.} Depending on where the VSI operates in a
disk, additional effects may become important. For example, in the
outer parts of a protoplanetary disk, where we estimate the VSI 
operates, disk self-gravity (SG) may play a role. Specifically, the vertical
shear profile depends on the vertical density stratification
(Eq. \ref{vertical_shear}) which can be enhanced by vertical SG. This
may promote the VSI. Inclusion of the Poisson equation should
therefore be considered.

@@

\emph{Radial dependence.} Based on previous works, including
non-linear simulations, which show that the VSI is relevant for
perturbations with small radial lengthscales, we have adopted the
radially-local approximation throughout this paper.  Although this
simplifies the linear problem to ordinary differential equations, this
approach introduces the wavenumber $k_x$ as a free parameter. The
effect of radial boundary conditions cannot be explored. Future 
generalizations of the linear problem should retain the radial
dependence properly and solve the two-dimensional $(r,z)$ partial 
differential equation eigenvalue problem.

>>>>

\subsection{Caveats and outlooks} 
We highlight below some extensions to the present  
linear models in future works.

\subsubsection{Parameter study and convective overstability}
We have focused on parameter regimes relevant to the VSI,
namely $\khat\gtrsim 1$ and $\beta\ll 1$. This excludes the
`convective overstability' recently discussed by \cite{klahr14} and
\cite{lyra14} in vertically unstratified disk models.

The convective overstability corresponds to growing epicyclic motions.
It is most effective for disturbances with $|k_x|\ll |k_z|$, where
$k_z$ is a vertical wavenumber, and $\beta\sim 1$. These conditions
are mutually exclusive with the VSI. 
The maximum growth 
rate for convective overstability is given by 
\begin{align}
  \nu_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
\end{align}
\citep{lyra14}, where $N_r^2 = -(c_p\rho)^{-1}\p_rP\p_rS$ is the
square of the radial buoyancy frequency
(cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.

Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
we find  
\begin{align}
  \nu_\mathrm{conv,max} =
  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
\end{align}
% We see that convective overstability is generally slower 
% than the VSI because $\nu_\mathrm{conv,max} = O(\epsilon^2\Omega_k)$,
% whereas the maximum VSI growth rate is $O(\epsilon\Omega_k)$. 
For the MMSN considered in this work (\S\ref{application} and Appendix
\ref{mmsn}), convective overstability does not apply because
$N_r^2>0$.

However, for more general disk structures convective overstability is
possible. In fact, the models adopted for our numerical study in
\S\ref{numerical} do satisfy $N_r^2<0$. %show early result
We will explore convective overstability more fully in a follow-up
study.  Although our linear code can be used for convective
overstability in principle, the fact convective overstability is
relevant for perturbations with $k_x\to0$ suggest radially global disk
models may be more appropriate.

% Nevertheless, it will be important to expand the parameter space covered in
% this study to include convective overstability,

%self-consistence may require radially global disk models

\subsubsection{Improved thermal evolution.} 
We assumed non-adiabatic effects
can be represented by a thermal relaxation timescale $t_c$. For 
application to protoplanetary disks we connected $t_c$ to the disk
properties, but one can improve the realism further by generalizing the
opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
\kappa_d(\rho, T)$. 
%two temperature
% structure

However, going beyond the framework of thermal
relaxation requires a proper treatment for radiative diffusion, which
involves the Lapacian operator and additional boundary conditions
must be supplemented. We will consider these improvements in a
follow-up study.

\subsubsection{Additional physics.}

Depending on where the VSI operates in a
disk, additional effects may become important. For example, in the
outer parts of a protoplanetary disk, where we estimate the VSI 
operates, disk self-gravity (SG) may play a role. Specifically, the vertical
shear profile depends on the vertical density stratification
(Eq. \ref{vertical_shear}) which can be enhanced by vertical SG. This
may promote the VSI. Inclusion of the Poisson equation should
therefore be considered.

@@

% \emph{Radial dependence.} Based on previous works, including
% non-linear simulations, which show that the VSI is relevant for
% perturbations with small radial lengthscales, we have adopted the
% radially-local approximation throughout this paper.  Although this
% simplifies the linear problem to ordinary differential equations, this
% approach introduces the wavenumber $k_x$ as a free parameter. The
% effect of radial boundary conditions cannot be explored. Future 
% generalizations of the linear problem should retain the radial
% dependence properly and solve the two-dimensional $(r,z)$ partial 
% differential equation eigenvalue problem.
